---
title: "Assignment38"
author: "John Bute"
date: "2024-11-19"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cran.rstudio.com/"))
```

## R Markdown

3. Construct and evaluate naïve Bayes classifiers for the Wine and for the 2011 Gapminder dataset.
```{r}
library(e1071)
Wine <- read.csv("C:/Users/johnb/Desktop/Machine Learning/data/wine.csv", stringsAsFactors = TRUE)

set.seed(1)
Wine$Class = as.factor(Wine$Class)

index <- sample(1:nrow(Wine), 0.7 * nrow(Wine))
train_data <- Wine[index, ]
test_data <- Wine[-index, ]

nb_model <- naiveBayes(Class ~ ., data = train_data)

summary(nb_model)
nb_preds <- predict(nb_model, newdata = test_data, type = "class")
conf_matrix_nb <- table(Actual = test_data$Class, Predicted = nb_preds)
print(conf_matrix_nb)

test_error <- 1 - sum(diag(conf_matrix_nb)) / sum(conf_matrix_nb)
cat("Test Error Rate:", test_error, "\n")
```
This model performs incredibly well in classifying all three wine classes across all metrics. It can predict classes 1 and 3 100% accurately, with a small mistep for class 2. 


4. Construct and evaluate CART models for the Wine and for the
Wisconsin Breast Cancer datasets.

```{r}
install.packages("rpart.plot")
library(rpart)
library(rpart.plot)  
set.seed(1)
cart_model <- rpart(Class ~ ., data = train_data, method = "class")
rpart.plot(cart_model)

summary(cart_model)
cart_preds <- predict(cart_model, newdata = test_data, type = "class")
conf_matrix_cart <- table(Actual = test_data$Class, Predicted = cart_preds)
print(conf_matrix_cart)

test_error <- 1 - sum(diag(conf_matrix_cart)) / sum(conf_matrix_cart)
cat("Test Error Rate:", test_error, "\n")


```

Similarly, the cart model performs well, as it is strong in identifying class 1 and 3 wines, and overall, has an error rate of 7% only. Furthermore, the tree structure provides interpretable splits based on Flavanoids, Total.phenols, and hue.
