---
title: "Assignment3"
author: "John Bute"
date: "2024-09-21"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 3

Before we begin, we must develop our code. My logic was that instead of
computing the Confidence, Interest, ... of every rule, why not create a
function (calculate_rule_metrics) that would a return a list of the
calculated metrics of rule A-\>B. Then, I created a small database of
the rules to study, looped through them, and appended the results to a
dataframe as seen below. From there, we can start evaluating each rule.

```{r cars}
Transactions <- read.csv("C:/Users/johnb/Desktop/Machine Learning/data/Transactions.csv")

library(dplyr)

calculate_rule_metrics <- function(data, antecedent_cols, consequent_col) {
  if (!all(c(antecedent_cols, consequent_col) %in% names(data))) {
    stop("One or more specified columns do not exist in the dataframe.")
  }
  
  N <- nrow(data)
  
  freq_A <- data %>%
    filter(Reduce(`&`, lapply(antecedent_cols, function(col) data[[col]] == 1))) %>%
    nrow()
  
  freq_B <- sum(data[[consequent_col]] == 1)
  
  freq_A_and_B <- data %>%
    filter(Reduce(`&`, lapply(antecedent_cols, function(col) data[[col]] == 1)) & data[[consequent_col]] == 1) %>%
    nrow()
  
  support <- freq_A_and_B / N
  confidence <- freq_A_and_B / freq_A
  interest <- confidence - (freq_B / N)
  lift <- (N^2 * support) / (freq_A * freq_B)
  conviction <- (1 - (freq_B / N)) / (1 - confidence)
  
  return(list(
    Support = support,
    Confidence = confidence,
    Interest = interest,
    Lift = lift,
    Conviction = conviction
  ))
}

results_df <- data.frame(
  Rule = character(),
  Support = numeric(),
  Confidence = numeric(),
  Interest = numeric(),
  Conviction = numeric(),
  stringsAsFactors = FALSE
)

rules <- list(
  list(antecedent = c("Red", "White"), consequent = "Green"),
  list(antecedent = c("Green"), consequent = "White"),
  list(antecedent = c("Red", "Green"), consequent = "White"),
  list(antecedent = c("Green"), consequent = "Red"),
  list(antecedent = c("Orange"), consequent = "Red"),
  list(antecedent = c("White", "Black"), consequent = "Yellow"),
  list(antecedent = c("Black"), consequent = "Green")
)

for (rule in rules){
  antecedent <- rule$antecedent
  consequent <- rule$consequent
  metrics <- calculate_rule_metrics(Transactions, antecedent, consequent)
  
  if(!is.null(metrics)){
    results_df <- rbind(
      results_df,
      data.frame(
        Rule = paste("{", paste(antecedent, collapse = ", "), "} -> {", consequent, "}"),
        Support = metrics$Support,
        Confidence = metrics$Confidence,
        Interest = metrics$Interest,
        Lift = metrics$Lift,
        Conviction = metrics$Conviction
      )
    )
  }
}
print(results_df)
```

Among the 7 rules evaluated we have the following observations:

-   Highest Lift: White and Black -\> Yellow
    -   Transactions with White and Black, that lead to yellow are
        positively correlated.
-   Highest Confidence: White and Black -\> Yellow
    -   If a customer were to buy white and black, then we can strongly
        assume that they will also buy Yellow
-   Highest Interest: White and Black -\> Yellow
    -   We can strongly assume that people who buy White and Black do
        not just randomly buy yellow too. Rather, it is a strong
        pattern.
-   Highest Conviction: White and Black -\> Yellow
    -   A high conviction states that this rule is stronger than
        expected by random chance.

We will now proceed to inventing randomly 7 new rules:

Black -\> Red \| Orange, White, Red -\> Green \| Green -\> Black
\| White -\> Black \| Red, Orange -\> Yellow \| Green, Black -\> White
\| Red -\> Black

Which we will now include in our analysis

```{r}
library(dplyr)
library(knitr)

calculate_rule_metrics <- function(data, antecedent_cols, consequent_col) {
  if (!all(c(antecedent_cols, consequent_col) %in% names(data))) {
    stop("One or more specified columns do not exist in the dataframe.")
  }
  
  N <- nrow(data)
  
  freq_A <- data %>%
    filter(Reduce(`&`, lapply(antecedent_cols, function(col) data[[col]] == 1))) %>%
    nrow()
  
  freq_B <- sum(data[[consequent_col]] == 1)
  
  # Calculate frequency of A and B occurring together
  freq_A_and_B <- data %>%
    filter(Reduce(`&`, lapply(antecedent_cols, function(col) data[[col]] == 1)) & data[[consequent_col]] == 1) %>%
    nrow()
  
  support <- freq_A_and_B / N
  confidence <- freq_A_and_B / freq_A
  interest <- confidence - (freq_B / N)
  lift <- (N^2 * support) / (freq_A * freq_B)
  conviction <- (1 - (freq_B / N)) / (1 - confidence)
  
  return(list(
    Support = support,
    Confidence = confidence,
    Interest = interest,
    Lift = lift,
    Conviction = conviction
  ))
}

results_df <- data.frame(
  Rule = character(),
  Support = numeric(),
  Confidence = numeric(),
  Interest = numeric(),
  Conviction = numeric(),
  stringsAsFactors = FALSE
)

rules <- list(
  list(antecedent = c("Red", "White"), consequent = "Green"),
  list(antecedent = c("Green"), consequent = "White"),
  list(antecedent = c("Red", "Green"), consequent = "White"),
  list(antecedent = c("Green"), consequent = "Red"),
  list(antecedent = c("Orange"), consequent = "Red"),
  list(antecedent = c("White", "Black"), consequent = "Yellow"),
  list(antecedent = c("Black"), consequent = "Green"),
  list(antecedent = c("Black"), consequent = "Red"),
  list(antecedent = c("Orange", "White", "Red"), consequent = "Green"),
  list(antecedent = c("Green"), consequent = "Black"),
  list(antecedent = c("White"), consequent = "Black"),
  list(antecedent = c("Red", "Orange"), consequent = "Yellow"),
  list(antecedent = c("Green", "Black"), consequent = "White"),
  list(antecedent = c("Red"), consequent = "Black")

)

for (rule in rules){
  antecedent <- rule$antecedent
  consequent <- rule$consequent
  metrics <- calculate_rule_metrics(Transactions, antecedent, consequent)
  
  if(!is.null(metrics)){
    results_df <- rbind(
      results_df,
      data.frame(
        Rule = paste("{", paste(antecedent, collapse = ", "), "} -> {", consequent, "}"),
        Support = metrics$Support,
        Confidence = metrics$Confidence,
        Interest = metrics$Interest,
        Lift = metrics$Lift,
        Conviction = metrics$Conviction
      )
    )
  }
}

print(results_df)

```

Out of these rules, there are 4 that seem extremly useful to managers.

Firstly , { Green } -\> { Black } is useful as its support and
confidence says that this pattern occurs very frequently.

Secondly, { Black } -\> { Green } is useful as its support and
confidence says that this pattern occurs very frequently.

Thirdly, { White, Black } -\> { Yellow } is extremly useful as its high
confidence, lift, and conviction mean that this is a highly reliable
pattern, where customers are likely to buy yellow items when buying
white and black ones. Furthermore, the lift suggests this is a strong
positive association.

Fourthly, { Red, Green } -\> { White } is useful due to its lift, as you
can use the rule to initiate bundling strategies.

NOTE: { Orange, White, Red } -\> { Green } has very strong metrics, as
its infinite conviction signifies that the action of buying green is
never absent when buying orange, white and red. However, it might not be
useful as this rule barely happens (0.0014 support value) and the
categories are almost the entire domain (4 out of the 6 colors).

To determine a threshold for support, coverage, interest, and lift of
rules derived from a given dataset, its important to know our dataset.
For support, if it was a big dataset, any threshold that is low (0.01)
is sufficient to take a look. However, in our case a 5% would be better
to avoid uninteresting rules. For confidence, the situation is
important, as for example, if you were using these rules to set up
targeted promotions, you would want a higher confidence. For lift, you
should look at all the rules you have discovered and determine what is a
good cutoff to filter out rules that do not seem impactful.

Finally, interest, anything above 0 ensures that rules are not
coincidental. Generally, you must first explore the situation, and then
the data, to make sure you fully comprehend what rule and its metrics
would be useful to you or the client in the situation. Finally,
benchmarking would be a good idea, as comparing the rules with known
patterns can ensure thresholds are not overly lenient or restrictive.
