---
title: "Assignment37"
author: "John Bute"
date: "2024-11-18"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
In this problem, you will use support vector approaches in order to predict whether a given car gets high or low gas mileage based on the Auto data set.


1. Create a binary variable that takes on a 1 for cars with gas mileage above the median, and a 0 for cars with gas mileage below the median.
```{r}
Auto <- read.csv("C:/Users/johnb/Desktop/Machine Learning/data/Auto.csv")
Auto <- na.omit(Auto)

median_mpg <- median(Auto$mpg)


Auto$mpg01 <- as.factor(ifelse(Auto$mpg > median_mpg, 1, 0))
Auto <- Auto[, !(names(Auto) %in% c("name", "mpg"))]


```

2. Fit a support vector classifier to the data with various values of cost, in order to predict whether a car
gets high or low gas mileage. Report the cross-validation errors associated with different values of this parameter. Comment on your results.
```{r}
library(e1071)

set.seed(1)
index <- sample(1:nrow(Auto), 0.7 * nrow(Auto))
train_data <- Auto[index, ]
test_data <- Auto[-index, ]

tune <- tune(
  svm,
  mpg01 ~ .,
  data = train_data,
  kernel = 'linear',
  ranges = list(cost = c(0.01, 0.05, 0.1, 0.5, 1, 1.5, 5, 10, 50, 100, 150))
)
summary(tune)

best_model <- tune$best.model

summary(best_model)

```
Our model has determined that the best one has C = 1.5



```{r}
test_preds <- predict(best_model, newdata = test_data, type = "class")
conf_matrix <- table(Actual = test_data$mpg01, Predicted = test_preds)
print(conf_matrix)

test_error <- 1 - sum(diag(conf_matrix)) / sum(conf_matrix)
cat("Test Error Rate:", test_error, "\n")
```
As we see, our svm performs pretty well, with an error rate of only 8%

3. Now repeat 2., this time using SVMs with radial and polynomial basis kernels, with different values of gamma and degree and cost. Comment on your results.
```{r}
tune_radial <- tune(
  svm,
  mpg01 ~ .,  
  data = train_data,
  kernel = "radial",
  ranges = list(
    cost =  c(0.01, 0.05, 0.1, 0.5, 1, 1.5, 5, 10, 50, 100, 150),
    gamma =  c(0.01, 0.05, 0.01, 0.1, 1, 10, 100 )
  )
)

summary(tune_radial)

best_model_radial <- tune_radial$best.model
print(best_model_radial)

```

As we see, with a radial model, our best performance come from a cost of 1 and a gamma of 1, with an error rate of 6.5%.

```{r}
test_preds_radial <- predict(best_model_radial, newdata = test_data, type = "class")
conf_matrix_radial <- table(Actual = test_data$mpg01, Predicted = test_preds_radial)
print(conf_matrix_radial)

test_error_radial <- 1 - sum(diag(conf_matrix_radial)) / sum(conf_matrix_radial)
cat("Test Error Rate:", test_error_radial, "\n")

```

```{r}
tune_poly <- tune(
  svm,
  mpg01 ~ .,  
  data = train_data,
  kernel = "polynomial",
  ranges = list(
    cost = c(0.01, 0.05, 0.1, 0.5, 1, 1.5, 5, 10, 50, 100, 150),
    degree = c(2:10)
  )
)
summary(tune_poly)
best_model_poly <- tune_poly$best.model
summary(best_model_poly)

test_preds_poly <- predict(best_model_poly, newdata = test_data, type = "class")

conf_matrix_poly <- table(Actual = test_data$mpg01, Predicted = test_preds_poly)
print(conf_matrix_poly)

test_error_poly <- 1 - sum(diag(conf_matrix_poly)) / sum(conf_matrix_poly)
cat("Polynomial SVM Test Error Rate:", test_error_poly, "\n")

```

For a polynomial kernel, we use degrees from 2 to 10 with the same cost vector. In total, our best cost is 5, with a degree of 3, and an error rate of 6.7%
4. Make some plots to back up your assertions in 2. and 3..

Hint: when p > 2, you can use the plot() function to create plots displaying pairs of variables at a time. Essentially, instead of typing plot(svmfit , dat) where svmfit contains your fitted model and dat is a data frame containing your data, you can type plot(svmfit , dat , x1~x4) in order to plot just the first and fourth variables. However, you must replace x1 and x4 with the correct variable names. To find out more, type ?plot.svm.

```{r}


plot(best_model, data = Auto, displacement ~ horsepower)

plot(best_model_radial, Auto, horsepower ~ weight)
plot(best_model_poly, Auto, horsepower ~ weight)
```


