---
title: "Q45"
author: "John Bute"
date: "2024-11-29"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

2. Cluster the Wine datasets using ğ‘˜âˆ’means, for various distance metrics and algorithm parameters. What is your best estimation for the number of
clusters in each case? Validate your results.

```{r}

data <- read.csv("C:/Users/johnb/Desktop/Machine Learning/data/wine.csv")
data <- scale(data)
```

```{r}

set.seed(1)

kmeans_clustering <- kmeans(data, centers = 3, nstart = 25)

print(kmeans_clustering)


library(factoextra)

fviz_nbclust(data, kmeans, method = "wss")


fviz_cluster(kmeans_clustering, data = data)

```

According to the elbow method, it appears to bend at 3 clusters, therefore 3 is the optimal number. 


3. Cluster the Wine datasets using hierarchical clustering, for various algorithm parameters. Validate your results.

```{r}

set.seed(1)

distance_matrix <- dist(data, method = "euclidean")

hc_clustering <- hclust(distance_matrix, method = "ward.D2")

plot(hc_clustering, main = "Hierarchical clustering", cex = 0.6)

hc_clusters <- cutree(hc_clustering, k = 3)


cluster_sizes <- table(hc_clusters)

print(cluster_sizes)

```

