---
title: "Q46"
author: "John Bute"
date: "2024-12-05"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Cluster the Wine datasets using DBSCAN, for various algorithm parameters. Validate your results.
```{r}

library(dbscan)
library(factoextra)

data <- read.csv("C:/Users/johnb/Desktop/Machine Learning/data/wine.csv")
data <- scale(data)


set.seed(1)
dbscan_cluster <- dbscan(data, eps = 3, minPts =10)

fviz_cluster(list(data = data, cluster = dbscan_cluster$cluster), geom = "point")+
  labs(title = "DBSCAN clustering")

print(table(dbscan_cluster$cluster))

```
Let us check the silhouette coefficient, to judge the quality of our clusters

```{r}
library(cluster)
silhouette_score <- silhouette(dbscan_cluster$cluster[dbscan_cluster$cluster != -1], dist(data[dbscan_cluster$cluster != -1,]))

avg_silhoueete_score <- mean(silhouette_score[, 3])

cat("Average silhouette score: ", avg_silhoueete_score, "\n")
```
Our silhouette score is weak. Furthermore, we do not have many dimensions, so our clustering does not suffer from dimensionality. Therefore, this clustering isn't great.

Cluster the Wine datasets using spectral clustering, for various algorithm parameters. Validate your results.
```{r}


library(kernlab)
library(cluster)
library(factoextra)

set.seed(1)

cluster_spectral <- specc(data, centers = 3)

fviz_cluster(list(data = data, cluster = as.integer(cluster_spectral))) +
  labs(title = "Spectral Clustering")

print(table(as.integer(cluster_spectral)))

```


We have three clusters now. Interesting

Lets check the silhouette score:

```{r}
cluster_spectral <- as.integer(cluster_spectral)
distance_matrix <- dist(data)
silhouette_score <- silhouette(cluster_spectral, distance_matrix)
avg_silhouette_score <- mean(silhouette_score[, 3])
cat("Average silhouette score: ", avg_silhouette_score, "\n")

```
We did way better than DBScan.
